Reasoning about a Capability Machine with Local Capabilities\\ Provably Safe Stack and Return Pointer Management (without OS Support)

### 2017-02-13 csf-2017/paper.tex
Two bibliography fixes.

The idea of applying SKLRs to an assembly language was applied before by
\citet{Hur:2011:KLR:1926385.1926402}, who constructed a binary SKLR for
formulating and proving correctness of an ML-to-assembly compiler. Their
condition on the assembly code are quite loose (for example, they allow
hand-written and self-modifying assmebly code), but they do assume code that it
is non-adversarial and respects the ABI of the high-level language (for example,
the code does not store return pointers and invoke them after their intended
lifetime). As a result, they can treat return pointers as evaluation contexts in
a lambda calculus and use biorthogonality~\cite{pitts_higher_1998}. This means
that they define the logical relation to accept evaluation contexts if they
produce related observations for related values and to accept terms if they
produce related observations when plugged into related evaluation contexts. In
our setting, there is no guarantee that adversarial code will treat return
pointers as intended, so, as described in Section~\ref{sec:logical-relation}, we
define our logical in a (arguably natural but still) novel way, using a single
orthogonal closure, i.e. accept terms if they produce related observations when
given related arguments, where those arguments include the return pointer.

Another novel aspect of our logical relation is how we model the temporary
nature of local capabilities. The main insight is that the temporary, revokable
nature of these capabilities can be seen as a generalisation of the special
status of evaluation contexts in lambda calculi. Closely related here is a study
by \citet{Dreyer:jfp12} (DNB) of the impact of higher-order state and control
effects on local relational reasoning. DNB formally capture this special status
of evaluation contexts using Kripke worlds with public and private future world
relations. Our public and private worlds are inspired by theirs and they are at
times remarkably similar. For example, the state-transitition system in the
proof of our assembly version of the awkward example, is exactly the same as the
one used by DNB for the original awkward example. Nevertheless, there are also
remarkable differences: in DNB's logical relation, public future worlds only
turn up in the continuation relation (which we do not have, see above). This is
different from our value relation which uses public and private future worlds
depending on the locality of the capability at hand. Additionally, our worlds
are a bit more complex because, to allow store-local capabilities, they can have
temporary regions that are only monotonous w.r.t. public future worlds, while
DNB's worlds are entirely permanent.

@@

%\bibliographystyle{IEEEtran}
\bibliographystyle{plain}
\bibliography{IEEEabrv,references}
\end{document}
\grid

>>>>

The idea of applying SKLRs to an assembly language was applied before by
\citet{Hur:2011:KLR:1926385.1926402}, who constructed a binary SKLR for
formulating and proving correctness of an ML-to-assembly compiler. Their
condition on the assembly code are quite loose (for example, they allow
hand-written and self-modifying assmebly code), but they do assume code that it
is non-adversarial and respects the ABI of the high-level language (for example,
the code does not store return pointers and invoke them after their intended
lifetime). As a result, they can treat return pointers as evaluation contexts in
a lambda calculus and use biorthogonality~\cite{pitts_operational_1998}. This
means that they define the logical relation to accept evaluation contexts if
they produce related observations for related values and to accept terms if they
produce related observations when plugged into related evaluation contexts. In
our setting, there is no guarantee that adversarial code will treat return
pointers as intended, so, as described in Section~\ref{sec:logical-relation}, we
define our logical in a (arguably natural but still) novel way, using a single
orthogonal closure, i.e. accept terms if they produce related observations when
given related arguments, where those arguments include the return pointer.

Another novel aspect of our logical relation is how we model the temporary
nature of local capabilities. The main insight is that the temporary, revokable
nature of these capabilities can be seen as a generalisation of the special
status of evaluation contexts in lambda calculi. Closely related to this is a
study by \citet{Dreyer:jfp12} (DNB) of the impact of higher-order state and
control effects on local relational reasoning. DNB formally capture this special
status of evaluation contexts using Kripke worlds with public and private future
world relations. Our public and private worlds are inspired by theirs and they
are at times remarkably similar. For example, the state-transitition system in
the proof of our assembly version of the awkward example, is exactly the same as
the one used by DNB for the original awkward example. Nevertheless, there are
also remarkable differences: in DNB's logical relation, public future worlds
only turn up in the continuation relation (which we do not have, see above).
This is different from our value relation which uses public and private future
worlds depending on the locality of the capability at hand. Additionally, our
worlds are a bit more complex because, to allow store-local capabilities, they
can have temporary regions that are only monotonous w.r.t. public future worlds,
while DNB's worlds are entirely permanent.

@@

%\bibliographystyle{IEEEtran}
\bibliographystyle{plainnat}
\bibliography{IEEEabrv,references}
\end{document}
\grid
